---
title       : Crime in Guatemala
author      : Ronaldo Echeverria Bardales
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax, quiz, bootstrap]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides

---

## Descripition

This is a small exploration of the crimes reported by the National Civil Police in the Republic of Guatemala, this information was taken from the site https://www.ine.gob.gt/ine/estadisticas/bases-de-datos/hechos-delictivos/

And can be viewed on the site https://ereb2002.shinyapps.io/crimeinguatemala/

The variables considered are:
* Incident department
* Description of the crime
* Crime date
* Time of crime

---  

## Visualization

You can display the information by dates.
```{r nvd3plot2, results = 'asis', comment = NA, message = F, echo = F}
library("dplyr")
require("rCharts")
library("tidyr")
library("stringr")
library("ggplot2")

setwd(".")

delitos <- read.csv("pnc_victimas_2016_2020.csv")
dic_delitos <- read.csv("delito_guatemala.csv")
dic_departamento <- read.csv("departamento_guatemala.csv")

delitos <- inner_join(delitos, dic_delitos)
delitos <- inner_join(delitos, dic_departamento)

delitos$fecha <- as.Date(paste(delitos$anio_ocu, paste(str_pad(delitos$mes_ocu,width=2,pad="0"), str_pad(delitos$día_ocu,width=2,pad="0"), sep="-"),sep="-"))
delitos$ocurrencia <- 1

tabla_dia <-     delitos %>% filter(Departamento == 'Guatemala') %>% 
      filter(Delito %in% c("Robo de vehículos","Robo de motocicletas")) %>% 
      filter(fecha >= as.Date('2016-01-01') & fecha <= as.Date('2020-12-31')) %>%
      group_by(Departamento, Delito, fecha) %>% summarise(Cantidad = sum(ocurrencia))

    #plot(tabla_dia$fecha, tabla_dia$Cantidad, col=tabla_dia$Delito)
  qplot(fecha, Cantidad, data=tabla_dia, colour = Delito)
    
```

---

## Visualization

You can display the information by hours.
```{r nvd3plot2.2, results = 'asis', comment = NA, message = F, echo = F, warning=F}
tabla_hora <-     delitos %>% filter(Departamento == 'Guatemala') %>% 
      filter(Delito %in% c("Robo de vehículos","Robo de motocicletas")) %>% 
      filter(fecha >= as.Date('2016-01-01') & fecha <= as.Date('2020-12-31')) %>%
      group_by(Departamento, Delito, hora_ocu) %>% summarise(Cantidad = sum(ocurrencia))

    #plot(tabla_dia$fecha, tabla_dia$Cantidad, col=tabla_dia$Delito)
  qplot(hora_ocu, Cantidad, data=tabla_hora, colour = Delito, xlim = c(0,24))
    
```

---

## Visualization

```{r nvd3plot2.3, results = 'asis', comment = NA, message = F, echo = F, warning=F}
tabla_hora <-     delitos %>% filter(Departamento == 'Guatemala') %>% 
      filter(Delito %in% c("Robo de vehículos")) %>% 
      filter(fecha >= as.Date('2016-01-01') & fecha <= as.Date('2020-12-31')) %>%
      group_by(Departamento, Delito, hora_ocu) %>% summarise(Cantidad = sum(ocurrencia))

prediccion <- lm(Cantidad ~ hora_ocu, data=tabla_hora)

plot(tabla_hora$hora_ocu, tabla_hora$Cantidad, xlim = c(0,24))
abline(prediccion)
```




